CDSR Human Data/Modelling Paper

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain "in the safe zone" during a military engagement.  The size, shape, and location of this area depend the the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions}(CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.  We evaluate two representation techniques against the human data and discuss their limitations as a cognitive systems solution to this problem.
\end{abstract}

@@

We begin by discussing the features of CDSRs using illustrative examples.  We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.  We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.  To explore these claims, we present the results of a human study, in which, people marked and rated points and regions for three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

\section{Context-Dependent Spatial Regions}
Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.  To identify these regions, one must identify the physical boundaries (e.g., walls), the objects within the room and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.

%Combine these figures into a single figure across
\begin{figure}
\centering
%% Change the classroom to be a picture stimuli
  \begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/classroom.png}
  \caption{The front of the classroom is the area between the desks and the whiteboard where people give presentations.}
  \end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
  \caption{Safety for the scout is defined by the location of the friendly turrets and the capabilities of the opposing force.}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
  \caption{A bottleneck occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
  \label{fig:safety}
  \end{subfigure}

@@

The long-term goal of this work is to create of agents that can naturally interact with people in human environments.  While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.  First, the representational approach should be transferable to new situations.  While humans do not demonstrate mastery after a single example of concept, they are able to begin using it immediately refining it over time.  Second, the representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.

@@

The study was designed to examine the following questions about spatial regions:
\begin{enumerate}
	\item Given a particular situation how well do people agree on the location and extent of a named spatial region?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named spatial region?
\end{enumerate}

To examine these questions we developed three different trials. In all three trial types the subject was presented with a visual stimuli (such as a room floor plan or a map) and an accompanying linguistic description (such as \emph{go to the front of the room}, however the way the task the subject was set varied across the trials:
\begin{enumerate}
	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the spatial region described in the description or to select a no region option.
	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.  
	\item Trial type 3: subjects were asked to mark on the visual stimuli, by clicking with the mouse, the location that they felt was most prototypical of the spatial description. 
\end{enumerate}

All three trial types use the same visual stimuli and accompanying spatial descriptions. Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a prototypical layout with the chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and a prototypical layout with the chairs far away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts. Conseqeuntly, there were 7 classroom stimuli (circle, U $\times 2$, close $\times 2$, far $\times 2$). In designing these stimuli we deliberately chose some layouts that were very prototypical or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}. By varying the prototypicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more prototypical. This would provide insight into research question 1, in particular it would point to the effect of prototypicality on the semantics of contextually defined spatial regions. Also, because the prototypicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.

@@

\subsection{Participants}

43 undergraduate students from two different third level institutes took part in the study.

@@

One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$. Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes. $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement. In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distribute across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used. Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as being poor to fair. This would indicate that there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli. In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli. It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses. Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can to interpreted as fair to good.

@@

The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, there results support the hypothesis that the defintion of spatial regions in a given context is effected by the location and orientation of objects in the context. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.

The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region. For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location. For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli. The selection of these points was informed by the results of a pilot study. Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain. Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain. In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g. $Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}). For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.

@@

\begin{table}
\centerline{
\begin{tabular}{l c c c l}
\hline
Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
\hline
Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\\\ % 2_5
\hline
Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
\end{table}

@@

-----------------

-----------------

Task 2: The participant selected a point denoting the sweetspot of the region
Analysis of results:
(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
We would hope that the variance of the subjects responses is lower than that of randomly generated points.
After the initial comparison with the random point sets we may exclude participants based on extreme variance.

(2) Does context matter?
Do the clusters of participants responses move?
- for each experimental stimulus compute the mean x and y location for the responses for that trial
- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.

-----------------

>>>>

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain "in the safe zone" during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions}(CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.  We discss the implications for these results on cognitive systems development.
\end{abstract}

@@

We begin by discussing the features of CDSRs using illustrative examples.  We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.  We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

\section{Context-Dependent Spatial Regions}
Consider the following regions (shown in Figure \ref{examples}): the front of a classroom , safety in military engagement, or the geographical feature of a bottleneck.  To identify these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract overtime with changes to the surrounding residents and businesses.

%Combine these figures into a single figure across
\begin{figure}
\centering
%% Change the classroom to be a picture stimuli
  \begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/classroom.png}
  \caption{\textit{The front of the classroom} could be defined as the area between the desks and the whiteboard where people give presentations.}
  \end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/safety-3.jpg}
  \caption{\textit{Safety for the scout} is contextually defined by the location of the friendly turrets and the capabilities of the opposing force.}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
  \includegraphics[width=\textwidth]{figures/bottleneck.JPG}
  \caption{A \textit{bottleneck} occurs when paths are confined to a narrow region by geographic features.  In this case, the water prevents ground movement.}
  \label{fig:safety}
  \end{subfigure}

@@

The long-term goal of this work is to create intelligent agents that can naturally interact with people in human environments.  While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.  First, the representational approach should be transferable to new situations.  While humans do not demonstrate mastery after a single example of a concept, they are able to begin using it immediately refining it over time.  Second, the representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.

@@

The study was designed to examine the following questions about spatial regions:
\begin{enumerate}
	\item Given a particular situation how well do people agree on the location and extent of a named context dependent spatial region?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named context dependent spatial region?
\end{enumerate}

To examine these questions, we developed a set of visual stimuli depicting a human environment (such as a room floor plan or a map) each paired with an accompanying linguistic description of a CDSR easily understood within the context of that environment (such as \emph{go to the front of the room}).  Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: (1) the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; (2) the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and (3) the military engagement stimuli were always accompanied by the description \emph{safety}.

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of chairs: a circular chair layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the layouts. Conseqeuntly, there were 7 classroom stimuli (circle, U $\times 2$, close $\times 2$, far $\times 2$). In designing these stimuli we deliberately chose some layouts that were very common or close variations on standard layouts - e.g., Figures \ref{fig:classroom-e} and \ref{fig:classroom-f} - and others were more novel - Figure \ref{fig:classroom-a} and \ref{fig:classroom-c}. By varying the typicality of the layouts we hoped to get some variation across the between subject agreement on a per stimuli basis, for example we expected to get a higher between subject agreement for the stimuli that were more common. This would provide insight into research question 1, in particular it would point to the effect of typicality on the semantics of contextually defined spatial regions. Also, because the typicality of the scenes was varied by adjusting the location and orientation of objects in the scenes, a variation in agreement across the stimuli would provide evidence that the orientation and location of objects do effect the location and extent of spatial regions.

@@

Using these stimuli pairs, we developed an online experiment with two different types of trials.
\begin{enumerate}
	\item Trial type 1: the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.  Subjects drew polygons by clicking to position the verticies with lines being automatically generated between each point indicated as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.      
	\item Trial type 2: a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.  The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region. In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}). One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped. In the classroom domain the expected sweetspot varied between the points depending on the stimuli. In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets. P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection. 
\end{enumerate}

\subsection{Participants and Procedure}

43 undergraduate students from two different third level institutes took part in the study. To control for sequence effects the presentation order of the trials were randomised for each participant as was the order of the stimuli presented within each trial.

@@

One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$. Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes. $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement. In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used. Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as being poor to fair. This would indicate that there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli. In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no regrion responses for each stimuli. It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses. Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can to interpreted as fair to good.

@@

The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.

The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region. For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests. There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. Table \ref{tab:flipped-point-likert-scores} lists the results of this trial. In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.

@@

\begin{table}
\centerline{
\begin{tabular}{l c c c l}
\hline
Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
\\
\hline
Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
\\
\hline
Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
\\
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
\end{table}

@@

Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.
